---
title: "The relationship between povery rate and prostate cancer rate varies between regions"
author: "Hanyu Zhong"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$ set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
```

```{r}
cancer <- read_csv("https://statsmaths.github.io/stat_data/cancer_inc_data.csv")
statesum <- group_summarize(cancer, state)
regionsum <- group_summarize(cancer, region)
```

###Thesis statement###
In the whole dataset of the United States, there is no relationship between the poverty rate and the percentage of people getting prostate cancer. However, The relationship varies between regions. The East North Central, Mountain, New England, Pacific and West North Central show a negative relationship between poverty rate and prostate cancer rate, and East South Central, Middle Atlantic, South Atlantic and West South Central show a positive relationship between the two variables. 

###Poverty and prostate cancer###
The graph below shows the relationship between poverty level and the incidence rate of prostate cancer. The observation points are separated by different colors into four equal categories according to poverty levels. Purple and blue colors mean relatively high poverty. Green and red colors mean relatively low poverty. The blue regression line indicates that there is no obvious relationship between the two variables, and the model below the graph confirms the relationship.  
```{r}
ggplot(cancer, aes (poverty, prostate)) +
  geom_point(aes(color = cut(poverty, 4)), size = 0.4) +
  geom_smooth(method = "lm")
```

The model here indicates no relationship between poverty and the percentage population getting prostate cancer in the U.S since the confidence level contains a 0.

```{r}
model <- lm_basic(poverty ~ 1 + prostate, data = cancer)
reg_table(model, level = 0.95)
```

###The contradiction###
However, the map below shows some disparities of the relationship between these two variables among different places. I summarized the data for each state. On the map below, each dot represents a state and its poverty level and percentage of prostate cancer. The bigger the dot, the higher the percentage of people getting prostate cancer. The lighter the color, the higher the level of poverty. From the map below, we can see that most of the light-colored dots (light green dots) are big, meaning that higher poverty region is likely to get prostate cancer. However, there are some contradictions. Some of the dark dots are big as well. This shows that the relationship between the poverty level and the percentage of people getting prostate cancer has disparities between regions, and can be influenced by another factor.

```{r}
qmplot(lon_mean, lat_mean, data = statesum, geom = "blank") +
  geom_point(aes(color = poverty_mean, size = prostate_mean), alpha = 0.8) +
  scale_color_viridis() 
```

###Regional difference of the relationships###
I separated the first line graph so it shows the disparities between regions. We can see that the East North Central, Mountain, New England, Pacific and West North Central all show a negative relationship between the two variables, and East South Central, Middle Atlantic, South Atlantic and West South Central show a positive relationship between the two variables. By comparing the graphs, I also found out that the more purple and blue points a graph have (the higher the poverty level as mentioned in the first graph), the graph is more likely to show a positive relationship. In other words, the higher the number of households in poverty in a region, the more positive the relationship between poverty level and the number of prostate cancer patients.

```{r}
ggplot(cancer, aes(poverty, prostate)) +
  geom_point(size = 0.1, aes(color = cut(poverty, 4))) +
  geom_smooth(method = "lm", size = 0.4) +
  facet_wrap(~region, scales = "free")
```

###Conclusions###
There is no relationship between poverty level and the incidence rate of prostate cancer from all the data collected in the U.S. However, there is a difference between the relationships among regions. In addition, the higher the number of households in poverty in that region, the more positive the relationship is. 

